---
title: "Getting started"
output:
  rmarkdown::html_vignette:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

## Introduction

{mlr3spatial} adds [mlr3::DataBackend]s for spatial classes ([terra::SpatRaster], [raster::brick], [stars::stars]).
The package is capable of making predictions to objects of these classes via `predict_raster()`.
The return is a spatial object in the class specified in argument `format`.

Essentially, {mlr3spatial} takes of the burden of converting spatial objects into a plain `data.table` and then coercing the predicted values back into the spatial object while making sure to not loose the spatial reference.

There are some more goodies in the bag though:

- Thanks to mlr3's ability to predict in parallel with any learner, {mlr3spatial} predictions can also make use of **future-based parallelization** and speed up the predictions of spatial objects.
Often enough, spatial predictions are quite large (in the millions of values) and efficient parallelization can save some time here.
See the vignette on ["Benchmarking parallel predictions"](https://mlr3spatial.mlr-org.com/articles/benchmark.html) for details.
- `predict_raster()` can be executed in chunks (argument `chunksize`), making it possible to **execute predictions on large raster files** on consumer grade machines.
  This chunked execution comes with some overhead compared to execution in one block since the prediction heuristic is executed multiple times and the results need to be merged in the end.

In the following, we showcase a step-by-step example how to handle a multi-layer raster object from package {stars}.

## Use Case - Landsat7 data as {stars} object

### Data Preparation

```{r}
library("mlr3")
library("mlr3spatial")
```

First, the TIFF files is read via `stars::read_stars()` and put into a `DataBackendRaster`.
The DataBackend is then used to create a regression task with the response being `band1`.

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
stack = stars::read_stars(tif)

backend = as_data_backend(stack)
task = as_task_regr(backend, target = "band1")

print(task)
```

For large raster files with millions of values it helps to predict in parallel.
To enable this, set `learner$parallel_predict = TRUE` and initiate a parallel plan via {future}.
Since this is only an example, parallelization is not enabled here.
Here we will use a simple regression tree as an example learner.
In practice you might want to use a different learner - you can find an overview of available learners [here](https://mlr3extralearners.mlr-org.com/articles/learners/list_learners.html).

```{r}
learner = lrn("regr.rpart")
set.seed(42)
row_ids = sample(1:task$nrow, 500)
learner$train(task, row_ids = row_ids)

print(learner)
```

### Prediction

For prediction `predict_spatial()` is used.
It will return a raster file which contains the predictions.
Users can select which R spatial format the returned raster should have.

In the following, we will compare the way to conduct the prediction using {mlr3spatial} with the "native" way of fitting an `e1071::svm()` model and predicting with `terra::predict()`.

#### mlr3spatial

```{r}
ras = predict_spatial(task, learner, format = "stars")
names(ras) = "cadmium"

print(ras)
```

#### stars

Since the layers are merged in a {stars} object, one first need to split them up and convert them into a regular data.table.
Next, the column names need to be adjusted to match the ones of the training data.
Afterwards, the `data.frame` generic of `predict()` can be called.
Finally, the predictions need to be injected into a stars object again.

(All of these steps are happening internally in {mlr3spatial}).

```{r}
rpart_learner = rpart::rpart(band1 ~ ., data = task$data(rows = row_ids))
stars_stack = as.data.table(split(stack, "band"))
stars_stack[, c("x", "y", "X1")] = NULL
colnames(stars_stack) = task$feature_names

stars_pred = predict(rpart_learner, stars_stack)

# subset stars object to one band only
stars_pred_ras = stack[, , , 1]
# rename the layer name
names(stars_pred_ras) = "pred"
# assign predictions
stars_pred_ras$pred = stars_pred

print(stars_pred_ras)
```

### Output consistency

Now that we have executed two predictions, we would like to verify that these are actually identical.

```{r}
all.equal(as.numeric(stars_pred_ras$pred), as.numeric(ras$cadmium))
```

### Visualization

Finally we can plot the predictions.
The color vector is extract from the viridis color palette via `dput(viridis::viridis_pal()(5))`.

```{r}
plot(ras, col = c("#440154FF", "#443A83FF", "#31688EFF", "#21908CFF", "#35B779FF", "#8FD744FF", "#FDE725FF"))
```
